VOI-aware Monte Carlo Sampling in Trees

### 2011-12-08 paper/vago/paper.tex
spelling edits

Following the principles of rational metareasoning, an arm with
the highest VOI should be pulled at each step. The upper bounds
established in Theorems~\ref{thm:vopi},~\ref{thm:be} can be used
as VOI estimates. Blinkered VOI estimates~(\ref{eqn:thm-be},
 \ref{eqn:bound-blnk-hoeffding}) can be viewed
as approximations of VOI estimates for myopic or
receding horison policies. As illustrated by the empirical evaluation
(Section~\ref{sec:empirical-evaluation}), estimates based on upper
bounds on the VOI result in rational sampling policies, and exhibit
performance comparable to the performance of some state-of-the-art
heuristic algorithms.

@@

One way to account for this re-use would be to incroprorate the
`future' value of information into a VOI estimate. However, this 
approach appears to be complicated. Alternatively, one can behave
myopically in search tree depth:
\begin{enumerate}
\item estimate VOI as though the information is discarded after each step;
\item stop early if the VOI is below a certain threshold
   (see Section~\ref{sec:control-termination-condition});
\item save the unused sample budget for search in future state, such that
   if the nominal budget is $N$, and the unused budget in the last state
   is $N_u$, the search budget in the next state will be $N+N_u$.
\end{enumerate}
In this approach, the cost of a sample in the current state is the
VOI of increasing the budget of a future state by one sample.  It is
unclear whether this cost can be accurately estimated, but supposedly
a fixed value for a given problem type and algorithm implementation
would work. Indeed, the empirical evaluation (Section~\ref{sec:emp-go})
confirms that stopping and sample redistribution based on a learned
fixed cost  substantially improve the performance of the VOI-based
sampling policy in game tree search.

@@

\begin{figure}[h]
\centering
\includegraphics[scale=0.8]{bests-bw.pdf}
\caption{Go: best winning rate comparison}
\label{fig:best-winning-rate}
\end{figure}
Figure~\ref{fig:best-winning-rate} presents relative performance of
different VOI-aware policies depending on the number of samples per
ply by comparing the best achieved winning rates against UCT (again,
the lower the bar, the better is the policy). Both ECT (based on the
Hoeffding inequality with midpoint) and BCT (based on the empirical
Benrstein inequality) are better than VCT. The difference between ECT
and VCT is most prominent for smaller number of samples, and the
winning rate of ECT increases relatively slowly with the number of
samples. In contrast, BCT is worse than ECT and only slightly better
than VCT for smaller number of samples, but the winning rate rapidly
increases with the number of samples, and for 10000 and 15000 samples
per ply BCT is the best of the policies with the winning rate reaching
67\%.

>>>>

Following the principles of rational metareasoning, an arm with
the highest VOI should be pulled at each step. The upper bounds
established in Theorems~\ref{thm:vopi},~\ref{thm:be} can be used
as VOI estimates. Blinkered VOI estimates~(\ref{eqn:thm-be},
 \ref{eqn:bound-blnk-hoeffding}) can be viewed
as approximations of VOI estimates for myopic or
receding horizon policies. As illustrated by the empirical evaluation
(Section~\ref{sec:empirical-evaluation}), estimates based on upper
bounds on the VOI result in rational sampling policies, and exhibit
performance comparable to the performance of some state-of-the-art
heuristic algorithms.

@@

One way to account for this re-use would be to incorporate the
`future' value of information into a VOI estimate. However, this 
approach appears to be complicated. Alternatively, one can behave
myopically in search tree depth:
\begin{enumerate}
\item estimate VOI as though the information is discarded after each step;
\item stop early if the VOI is below a certain threshold
   (see Section~\ref{sec:control-termination-condition});
\item save the unused sample budget for search in future state, such that
   if the nominal budget is $N$, and the unused budget in the last state
   is $N_u$, the search budget in the next state will be $N+N_u$.
\end{enumerate}
In this approach, the cost of a sample in the current state is the
VOI of increasing the budget of a future state by one sample.  It is
unclear whether this cost can be accurately estimated, but supposedly
a fixed value for a given problem type and algorithm implementation
would work. Indeed, the empirical evaluation (Section~\ref{sec:emp-go})
confirms that stopping and sample redistribution based on a learned
fixed cost  substantially improve the performance of the VOI-based
sampling policy in game tree search.

@@

\begin{figure}[h]
\centering
\includegraphics[scale=0.8]{bests-bw.pdf}
\caption{Go: best winning rate comparison}
\label{fig:best-winning-rate}
\end{figure}
Figure~\ref{fig:best-winning-rate} presents relative performance of
different VOI-aware policies depending on the number of samples per
ply by comparing the best achieved winning rates against UCT (again,
the lower the bar, the better is the policy). Both ECT (based on the
Hoeffding inequality with midpoint) and BCT (based on the empirical
Bernstein inequality) are better than VCT. The difference between ECT
and VCT is most prominent for smaller number of samples, and the
winning rate of ECT increases relatively slowly with the number of
samples. In contrast, BCT is worse than ECT and only slightly better
than VCT for smaller number of samples, but the winning rate rapidly
increases with the number of samples, and for 10000 and 15000 samples
per ply BCT is the best of the policies with the winning rate reaching
67\%.
